---
title: "User demo"
output:
  html_document:
    toc: true
    toc-location: right
    number_sections: false
    code_folding: show
---

## Hi, there!  
This is a demo for all the user inputs you can edit to run the various models and simulations.<br>
Inside we drafted different sections for each scenario (A to H). 

**Sections**
: A) Deterministic (constant beta)
: B) Deterministic (seasonal beta)
: C) Stochastic (constant beta)
: D) Stochastic (seasonal beta)
: E) Multi-pop deterministic
: F) Multi-pop stochastic
: G) Dashboards with self arrangement
: H) Summaries & helpers functions

Explore the code chunks below and edit the inputs! Have fun :)<br>

For installation please check the `README.md` instructions.<br>

## Edit your inputs here
```{r}
# ---- EDIT YOUR INPUTS HERE -------------------
# Common components
n_times <- 60      # Total days to simulate
pop     <- 100000  # Population size
I_init  <- 10      # Initial number of infectious individuals
beta    <- 0.75    # Transmission rate  
gamma   <- 1/7     # Recovery rate
omega   <- 1/30    # Rate of loss of immunity(R -> S)


# Seasonal beta (for B, D)
season_base  <- 0.70   # Base transmission rate
season_amp   <- 0.25   # Amplitude of seasonal variation (0 ≤ amplitude < 1)
season_phase <- 30     # Phase shift in days

# Stochastic (C,D,F)
n_sims   <- 200   # Number of independent simulation runs (columns).
epsilon  <- 1e-4  # External infection pressure.
alpha    <- 0.3   # Reporting rate
seed     <- 42    # Random seed for reproducibility

# Multi-pop (E, F)
mp_n_times <- 180                     # Total days to simulate
mp_pops    <- c(100000, 500, 200000)  # Population sizes
mp_Iinit   <- c(100, 5, 1000)         # Initial infectious individuals
mp_beta    <- 0.20                    # Transmission rate
mp_gamma   <- 1/7                     # Recovery rate
mp_omega   <- 1/30                    # Rate of loss of immunity(R -> S)

# Multi-pop stochastic (F)
mp_sims    <- 200      # Number of stochastic simulations
mp_epsilon <- 0        # Noise parameter
mp_alpha   <- NULL     # Scaling factor for stochasticity
mp_seed    <- 99       # Random seed for reproducibility

# Customise uncertainty ribbons(Default to 95%)
ribbon_probs <- c(0.025, 0.975)

# Customise colors for multiple populations
my_pop_cols <- c("#E41A1C", "#377EB8", "#4DAF4A")
```

## Load package functions
```{r}
# Please ensure you have "R" file in your working directory
R.utils::sourceDirectory("R/")
```

## A - Deterministic model with constant beta
```{r A-Deterministic}
simA <- simulate_sirs_det(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta, gamma = gamma, omega = omega
)

# plot options: "overlay", "sir", "incidence", "both_side"
plotA <- plot_sirs(simA, which = "overlay", per_million = FALSE)

print(plotA)
# If you want to save the plot, please use ggsave()
# ggsave(plotA, filename = "deterministic_constant_beta.png", width = 8, height = 5)
```

## B - Deterministic model with seasonal beta
```{r}
beta_B <- make_beta(
  n_times, mode = "seasonal",
  base = season_base, amplitude = season_amp, phase = season_phase
)
simB <- simulate_sirs_det(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta_B, gamma = gamma, omega = omega
)
# plot options: "overlay", "sir", "incidence", "both"
plot_sirs(simB, which = "sir")
```

## C - Stochastic model with constant beta(Single population)
```{r}
stoch_C <- simulate_sirs_stoch(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta, gamma = gamma, omega = omega,
  epsilon = epsilon, alpha = alpha,
  n_sims = n_sims, seed = seed
)
stoch_C$params$ribbon_probs <- ribbon_probs

# plot options: "overlay", "SIR", "incidence", "both"
plot_stoch(stoch_C, which = "SIR")
```

## D - Stochastic model with seasonal beta(Single population)
```{r}
betaD <- make_beta(
  n_times, mode = "seasonal",
  base = season_base, amplitude = season_amp, phase = season_phase
)
stochD <- simulate_sirs_stoch(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = betaD, gamma = gamma, omega = omega,
  epsilon = epsilon, alpha = alpha,
  n_sims = n_sims, seed = seed
)
# prefer explicit vector; plot_stoch should read this if implemented
stoch_C$params$ribbon_probs <- ribbon_probs

# plot options: "overlay", "SIR", "incidence", "both"
plot_stoch(stochD, which = "incidence")   # per_million=FALSE by default
```

## E - Multi-population deterministic model
```{r}
beta_md <- matrix(mp_beta, nrow = mp_n_times, ncol = length(mp_pops))
simE <- simulate_sirs_multi(
  n_times = mp_n_times, pop_vec = mp_pops, I_init = mp_Iinit,
  beta_mat = beta_md, gamma = mp_gamma, omega = mp_omega
)

# plot options: "S","I", "R", "incidence"
# group_style: "combined", "facet"
plot_multi(simE, which = "S", group_style = "combined")
```

## F - Multi-population stochastic model
```{r}
simF <- simulate_sirs_multi_stoch(
  n_times    = mp_n_times,
  pop_vec    = mp_pops,
  I_init     = mp_Iinit,
  beta_mat   = mp_beta,     
  gamma      = mp_gamma,
  omega      = mp_omega,
  epsilon    = mp_epsilon,
  alpha      = mp_alpha,
  n_sims     = mp_sims,
  seed       = mp_seed
)
simF$params$ribbon_probs <- ribbon_probs

# Plotting 
plot_multi(simF,                       # Change your data frame here
           which = "I",                # plot options: "S","I", "R", "incidence"
           group_style = "combined",   # group_style: "combined", "facet"
           show_bands = TRUE,          # show uncertainty ribbons or not
           per_million = TRUE)        # show per million or not(for incidence)
```

## H - Dashboards with self arrangement
```{r warning=FALSE, message=FALSE}
plot_DB <- plot_dashboard(
  sim           = simF,           # Try simD for single-pop stochastic
  probs         = c(0.95),        # Uncertainty ribbons
  per_million   = TRUE,           # per_million: TRUE, FALSE
  group_style   = "combined",     # group_style: "combined", "facet"
  pop_cols      = my_pop_cols,    # Custom population colors
  show_bands    = FALSE           # Stochastic ribbons visible or not
)

# plot options: "SIR", S","I", "R", "incidence","beta","params"
panels_DB <- c("S","I","R","incidence")

# Arrange dashboard
p_DB <- arrange_dashboard(plot_DB[panels_DB],     # Select panels
                          layout = c(2,2),        # Change layout here(rows, cols)
                          collect_legend = TRUE)  # Collect legend in one panel

print(p_DB)
```

## I - Summaries & helpers functions

### `summarize_sim` summarize simulation output.
```{r}
# Change different data frames here to see different summaries
sumF <- summarize_sim(simF, level = "per_pop")
```

```{r}
library(DT)

datatable(
  sumF,                             # Change the data frame here
  rownames = FALSE,
  filter = "top",                   # adds filter boxes to each column
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 25, 50, 100),
    autoWidth  = TRUE
  ),
  class = "compact stripe hover"
)

```

### `to_tidy()`Convert simulation output to a tidy long table.
```{r}
# Example of Multi-population stochastic model
tt <- to_tidy(simA)   # Change the data frame here
```

```{r}
# Filter for specific group and state
tt <- to_tidy(simF)|> 
  dplyr::filter(group == 1, state == "I")

# The table is large; show first 10 rows
head(tt,10)
```

### `make_beta()`Create time-varying beta (constant, seasonal).
```{r}
beta_seasonal <- make_beta(
  n_times = 365,        # Total days
  mode = "seasonal",    # Options: "constant", "seasonal"
  base = 0.7,           # Base transmission rate
  amplitude = 0.25,     # Amplitude of seasonal variation
  phase = 30            # Phase shift in days
)

ggplot(data = data.frame(day = 1:365, beta = beta_seasonal), aes(x = day, y = beta)) +
geom_line() +
labs(title = "Seasonal Beta over Time", x = "Day", y = "Beta") +
theme_minimal()
```

### `adjust_beta()` Scale beta within specified day windows
```{r}
B <- adjust_beta(
rep(0.2, 365), # Initial beta vector
data.frame(start = 50, end = 80), # Day windows to adjust
0.7 # Scaling factor
)

ggplot(data = data.frame(day = 1:365, beta = B), aes(x = day, y = beta)) +
geom_line() +
labs(title = "Beta over Time", x = "Day", y = "Beta") +
theme_minimal()
```

### `cumulative_incidence()` Calculate cumulative incidence over time

```{r}
ci_F <- cumulative_incidence(simF)

# The table is large; show first 10 rows

head(ci_F, 10)
```

### `attack_rate()` Calculate attack rate from simulation output

```{r}
# Attack rate (single-pop stochastic): vector length = n_sims
ar_C <- attack_rate(stoch_C)
head(ar_C)

# Attack rate (multi-pop stochastic): matrix [n_sims x n_groups]

ar_F <- attack_rate(simF)
dim(ar_F)
ar_F[1:5, ]
```

### `reff_from_sim()` Calculate effective reproduction number from simulation output

```{r}
# Reff from simulated paths
reff_C <- reff_from_sim(stoch_C)
head(reff_C)

# Reff helper (deterministic multi-pop)
reff_E <- reff_from_sim(simE)
head(reff_E)
```

If any questions or issues arise, please open an issue on the GitHub repository or reach out to the maintainers. Happy modeling!