---
title: "User demo"
output:
  html_document:
    toc: true
    toc-location: right
    number_sections: false
    code_folding: show
---

Hi, there!  
This is a demo for all the user inputs you can edit to run the various models and simulations.
Inside we drafted different sections for each scenario (A to H).

**Sections**
: A) Deterministic (constant beta)
: B) Deterministic (seasonal beta)
: C) Stochastic (constant beta)
: D) Stochastic (seasonal beta)
: E) Multi-pop deterministic
: F) Multi-pop stochastic
: G) Summary table of simulation results
: H) Dashboards with self arrangement

## Edit your inputs here
```{r}
# ---- EDIT YOUR INPUTS HERE -------------------
# Common components
n_times <- 60      # Total days to simulate
pop     <- 100000  # Population size
I_init  <- 10      # Initial number of infectious individuals
beta    <- 0.75    # Transmission rate  
gamma   <- 1/7     # Recovery rate
omega   <- 1/30    # Rate of loss of immunity(R -> S)


# Seasonal beta (for B, D)
season_base  <- 0.70   # Base transmission rate
season_amp   <- 0.25   # Amplitude of seasonal variation (0 ≤ amplitude < 1)
season_phase <- 30     # Phase shift in days

# Stochastic (C,D,F)
n_sims   <- 200   # Number of independent simulation runs (columns).
epsilon  <- 1e-4  # External infection pressure.
alpha    <- 0.3   # Reporting rate
seed     <- 42    # Random seed for reproducibility

# Multi-pop (E, F)
mp_n_times <- 180                     # Total days to simulate
mp_pops    <- c(100000, 500, 200000)  # Population sizes
mp_Iinit   <- c(100, 5, 1000)         # Initial infectious individuals
mp_beta    <- 0.20                    # Transmission rate
mp_gamma   <- 1/7                     # Recovery rate
mp_omega   <- 1/30                    # Rate of loss of immunity(R -> S)

# Multi-pop stochastic (F)
mp_sims    <- 200      # Number of stochastic simulations
mp_epsilon <- 0        # Noise parameter
mp_alpha   <- NULL     # Scaling factor for stochasticity
mp_seed    <- 99       # Random seed for reproducibility

# Customise uncertainty ribbons(Default to 95%)
ribbon_probs <- c(0.025, 0.975)

# Customise colors for multiple populations
my_pop_cols <- c("#E41A1C", "#377EB8", "#4DAF4A")
```

## Load the package
```{r}
# library(epi-simulation)
R.utils::sourceDirectory("R/")
```

## A - Deterministic model with constant beta
```{r A-Deterministic}
simA <- simulate_sirs_det(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta, gamma = gamma, omega = omega
)

# plot options: "overlay", "sir", "incidence", "both_side"
plot_sir_diag(simA, which = "overlay")
```

## B - Deterministic model with seasonal beta
```{r}
beta_B <- make_beta(
  n_times, mode = "seasonal",
  base = season_base, amplitude = season_amp, phase = season_phase
)
simB <- simulate_sirs_det(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta_B, gamma = gamma, omega = omega
)
# plot options: "overlay", "sir", "incidence", "both_side"
plot_sir_diag(simB, which = "sir")
```

## C - Stochastic model with constant beta(Single population)
```{r}
stoch_C <- simulate_sirs_stoch(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = beta, gamma = gamma, omega = omega,
  epsilon = epsilon, alpha = alpha,
  n_sims = n_sims, stochastic = TRUE, seed = seed + 1
)
stoch_C$params$ribbon_probs <- ribbon_probs

# plot options: "overlay", "SIR", "incidence", "both_side"
plot_stoch(stoch_C, which = "SIR")
```

## D - Stochastic model with seasonal beta(Single population)
```{r}
betaD <- make_beta(
  n_times, mode = "seasonal",
  base = season_base, amplitude = season_amp, phase = season_phase
)
stochD <- simulate_sirs_stoch(
  n_times = n_times, pop = pop, I_init = I_init,
  beta = betaD, gamma = gamma, omega = omega,
  epsilon = epsilon, alpha = alpha,
  n_sims = n_sims, stochastic = TRUE, seed = seed
)
# prefer explicit vector; plot_stoch should read this if implemented
stoch_C$params$ribbon_probs <- ribbon_probs

# plot options: "overlay", "SIR", "incidence", "both_side"
plot_stoch(stochD, which = "incidence")   # per_million=FALSE by default
```

## E - Multi-population deterministic model
```{r}
beta_md <- matrix(mp_beta, nrow = mp_n_times, ncol = length(mp_pops))
simE <- simulate_sirs_multi(
  n_times = mp_n_times, pop_vec = mp_pops, I_init = mp_Iinit,
  beta_mat = beta_md, gamma = mp_gamma, omega = mp_omega
)

# plot options: "S","I", "R", "incidence"
# group_style: "combined", "facet"
plot_multi(simE, which = "S", group_style = "combined")
plot_multi(simE, which = "I", group_style = "facet")
```

## F - Multi-population stochastic model
```{r}
simF <- simulate_sirs_multi_stoch(
  n_times    = mp_n_times,
  pop_vec    = mp_pops,
  I_init     = mp_Iinit,
  beta_mat   = mp_beta,     # scalar OK; recycled internally
  gamma      = mp_gamma,
  omega      = mp_omega,
  epsilon    = mp_epsilon,
  alpha      = mp_alpha,
  n_sims     = mp_sims,
  stochastic = TRUE,
  seed       = mp_seed
)
simF$params$ribbon_probs <- ribbon_probs

# plot options: "S","I", "R", "incidence"
# group_style: "combined", "facet"
plot_multi(simF, which = "I", group_style = "combined")
```

## G - Summary table of simulation results
```{r}
# Change different data frames here to see different summaries
sumF <- summarize_sim(simF, level = "per_pop")
```

```{r}
library(DT)

datatable(
  sumF,                             # Change the data frame here
  rownames = FALSE,
  filter = "top",                   # adds filter boxes to each column
  options = list(
    pageLength = 10,
    lengthMenu = c(10, 25, 50, 100),
    autoWidth  = TRUE
  ),
  class = "compact stripe hover"
)

```

## H - Dashboards with self arrangement
```{r warning=FALSE, message=FALSE}
plot_DB <- plot_dashboard(
  sim           = simF,           # Try simD for single-pop stochastic
  probs         = c(0.95),        # Uncertainty ribbons
  per_million   = TRUE,           # per_million: TRUE, FALSE
  group_style   = "combined",     # group_style: "combined", "facet"
  pop_cols      = my_pop_cols,    # Custom population colors
  show_bands    = FALSE           # Stochastic ribbons visible or not
)

# plot options: "SIR", S","I", "R", "incidence","beta","params"
panels_DB <- c("S","I","R","incidence")

# Arrange dashboard(Change layout as needed)
p_DB <- arrange_dashboard(plot_DB[panels_DB], layout = c(2,2), collect_legend = TRUE)

print(p_DB)
```

